Supplementary Informations to Ion pairing controls rheological properties of ``crocodile-line'' polyelectrolyte hydrogels

### 2016-10-23 response.tex

First we would like to thank the reviewer for having carefully read our manuscript and provided valuable comments and suggestions.

\begin{quotationi}
The authors report on rheological characterization and theoretical interpretation of properties of polyelectrolyte assemblies prepared from molecules with a single anionic headgroup and cationic groups along the chain. The presented work including preparation of the polyelectrolytes, their rheological characterization in solution as well as theoretical interpretation, is considered to be conducted in a nice way, it indicate the role of counterions in affecting the assembly modes of the structures and novel assembly modes are suggested. There are, however, some concerns that should be addressed before publication can be recommended:

Although offering novel handles to generate molecular assemblies, and for this reason, is considered to be general interest, the suggestion for the particular structures being assembled could be further bolstered by scattering tools.
\end{quotationi}

In many instances we tried to perform DLS measurements on our samples and, indeed, signals were observed. However the obtained results were largely unreproducible (even for a given sample upon slight change of the concentration, not to mention when two different samples were used). An example is shown below.

\includegraphics[width=\textwidth]{DLS.png}
\textit{Figure: \textbf{a} DLS data and best fit \textbf{b} size distribution by diffusion intensity and \textbf{c} size distribution by number for PImBr (top row) and PPyBr (bottom row) in water at 10\% weight concentrations}

As the Reviewer can see, dispersion was very high, and signal quality so low that cumulant fit analyses were absolutely irrelevant. In our opinion, there is nothing relevant in these data, this is why we finally chose not to include it in the SI. If the referee thinks that it might be worth adding some of these data to the article, we are ready to do it. However, we do know whether a reasonable interpretation could be provided from the data. We leave the referee judge; top is PImBr bottom PPyBr. Although a pattern seems to exist, it seems difficult to us to get its exact meaning, and to separate the artefact from the fact\ldots

@@

\begin{quotationi}
In applying network theory to interpret the data, it appears that e.g. loose end corrections are not employed. What is the rational for that?
\end{quotationi}

To our understanding, loose end corrections describe a situation where pre-existing linear chains of size $M$ are randomly cross-linked with an average of $M_c$ monomers between cross links with $M_c\leq M/2$. The modulus then scales as
\begin{equation}
\frac{G}{k_\mathrm{B}T} = \frac{c}{M_c}  \left(1 - \frac{2M_c}{M}\right).
\end{equation}
For example if $M_c$ is only one third of $M$, then two third of the original linear chain are loose ends and do not contribute to the elasticity.

This is rather different from our situation where the number of monomers between cross links is larger or equal to the number of monomers in the original chain. This situation stems from the reactivity of the heads: a cross link necessarily occurs at the head of a polymer, preventing one of the possible loose end, but leaving the possibility of many loose ends at the tails of each polymer.

We can better compare our situation to a gel formed by branched polymers, for example bottle brush polymers. In this situation, Papagiannopoulos et al. (\href{http://doi.org/10.1021/bm060287d}{Biomacromolecules 2006} or \href{http://doi.org/10.1039/b714864j}{Faraday Discussions 2008}) have shown that the shear modulus scales as
\begin{equation}
\frac{G}{k_\mathrm{B}T} = \frac{c}{N_\mathrm{tot}}.
\end{equation}
where $N_\mathrm{tot}$ is the total number of monomers between cross-links, including side chains that are topological `loose ends'.

In other words, one can consider two possible definitions of ``cross link''. In Flory's definition, every branching point is a cross link, therefore a brush has many loose ends. In the more coarse-grained definition used by Papagiannopoulos et al., branching points leading to a loose end are not considered meaningful cross links, but every monomer of the side chains actually contribute to entropy and thus to elasticity.

### 2016-10-24 PALSE/cover_letter.tex
Final version of the cover letter (forgot to commit)

Please find enclosed our manuscript, entitled \textsc{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}, which we are submitting for publication as an article to \journal.

Polymer-based hydrogels have promising biomedical applications, eg. as a soft scaffold for cellular growth. To be used in vivo, solid pieces of gel have to be implanted surgically. By contrast a physical gel with a yield stress rheology would allow a a much less traumatic delivery via injection: solid in the syringe, the gel could flow through the needle and reform as a solid once in the body. In the present work we design and study such a polymeric physical gel.

The system we investigate is made of short (70 units) linear polymer chains. In aqueous solution, a single extremity (the `head') is charged negatively whereas the repeated unit of the `body' carries a positive charge. We demonstrate that these chains self assemble into long processions (up to 800 chains) due to electrostatic interactions. %These processions are entangled or cross-linked on the scale of the screening length. 
The resulting physical hydrogel behaves as a yield stress fluid. In this article we build a microscopic understanding of the chemical physics involved in this gel that allows us engineer its rheological properties.

@@

Our findings are at the interface between polymer chemistry and rheology, bridged by an understanding of the underlying physics, with possible biomedical applications.

@@

We suggest the following scientists as possible reviewers, by alphabetical order:
\begin{enumerate} \setlength\itemsep{0em}
\item Andrey V. Dobrynin, University of Connecticut, 
avd@ims.uconn.edu
\item Elie Raphaël, UMR CNRS Gulliver 7083, ESPCI, elie.raphael@espci.fr
\item Michael Rubinstein, University of North Carolina, mr@unc.edu
\item Annette M. Schmidt, University of Cologne, annette.schmidt@uni-koeln.de
\item Nadine Schwierz, University of California, Berkeley, nschwierz@berkeley.edu
\end{enumerate}

>>>>

Please find enclosed our manuscript, entitled \textsc{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}, which we are submitting for publication as a full paper to \journal.

Polymer-based hydrogels have promising biomedical applications, eg. as a soft scaffold for cellular growth. To be used in vivo, solid pieces of gel have to be implanted surgically. By contrast a physical gel with a yield stress rheology would allow a a much less traumatic delivery via injection: solid in the syringe, the gel could flow through the needle and reform as a solid once in the body. In the present work we design and study a family of such polymeric physical gels.

The systems we investigate are made of short (70 units) linear polymer chains. In aqueous solution, a single extremity (the `head') is charged negatively whereas the repeated unit of the `body' carries a positive charge. We demonstrate that these chains self assemble into long processions (up to 800 chains) due to electrostatic interactions. %These processions are entangled or cross-linked on the scale of the screening length. 
The resulting physical hydrogel behaves as a yield stress fluid. In this article we build a microscopic understanding of the chemical physics involved in this gel that allows us engineer its rheological properties.

@@

Our findings are at the interface between polymer chemistry and rheology, bridged by an understanding of the underlying physics, with possible biomedical applications. In our opinion they are relevant and of importance for a broad community of scientists that encompass polymer and material chemists, physicochemists and soft matter physicists.

@@

We suggest the following scientists as possible reviewers, by alphabetical order:
\begin{enumerate} \setlength\itemsep{0em}
\item Andrey V. Dobrynin, University of Connecticut, 
avd@ims.uconn.edu
\item Kwang S. Kim, Ulsan National Institute of Science and Technology, kimks@unist.ac.kr
\item Elie Raphaël, UMR CNRS Gulliver 7083, ESPCI, elie.raphael@espci.fr
%\item Michael Rubinstein, University of North Carolina, mr@unc.edu
\item Annette M. Schmidt, University of Cologne, annette.schmidt@uni-koeln.de
\item Nadine Schwierz, University of California, Berkeley, nschwierz@berkeley.edu
%\item István Szilágyi, University of Geneva, Istvan.Szilagyi@unige.ch
\end{enumerate}
